---
title: "r2readthedocs"
output:
  md_document:
    variant: gfm

  rmarkdown::html_vignette:
    self_contained: no
---

<!-- badges: start -->
[![R build status](https://github.com/ropenscilabs/r2readthedocs/workflows/R-CMD-check/badge.svg)](https://github.com/ropenscilabs/r2readthedocs/actions)
[![codecov](https://codecov.io/gh/ropenscilabs/r2readthedocs/branch/main/graph/badge.svg)](https://codecov.io/gh/ropenscilabs/r2readthedocs)
[![Project Status:
Concept](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
<!-- badges: end -->

# r2readthedocs

Convert R package documentation to a ['readthedocs'](https://readthedocs.org/) website.

## Why?

Two compelling reasons:

1. ['readthedocs'](https://readthedocs.org/) is based on a markup language
   called ['reStructuredText'
   (`.rst`)](https://www.python.org/dev/peps/pep-0287/#benefits) which is more
   flexible and powerful than the `markdown` generally used in R packages. 
2. The structure of an R package documentation website generated with existing
   tools such as [`pkgdown`](https://pkgdown.r-lib.org/) is generally fixed and
   difficult to modify. The entire structure of
   a ['readthedocs'](https://readthedocs.org/) site can be very easily modified
   through changes to the underlying `.rst` files, including easy addition of
   documentation beyond the internal package-specific documentation, such as
   descriptions of use cases, of relevant data or analyses external to the
   package, or anything at all.

## How?

For those unfamiliar with ['readthedocs'](https://readthedocs.org/) (RTD), we
recommend using the `rtd_dummy_pkg()` function provided with this package,
which generates a "skeleton" of a package which can be used to generate an RTD
site. 

```{r rtd-dummy, eval = FALSE}
path <- rtd_dummy_pkg ()
```

Then either using that `path` to the dummy package, or a local path to a
package of your choice, generate an RTD website by running:

```{r rtd-run, eval = FALSE}
r2readthedocs (path)
```

The RTD website will be automatically opened in your default browser. Most of
the content is automatically generated straight from the package documentation,
as for a [`pkgdown`](https://pkgdown.r-lib.org/) website. The primary
difference you'll immediate notice is that the front page is not the package's
`README.md` document. This is because, which `markdown` pages can be readily
inserted into `.rst` pages, this is not possible for the main page, which must
be encoded entirely in `.rst` form. The front page of any site generated with
the `r2readthedocs()` function will retain this general form, which can then be
readily adapted as desired. The main file controlling the site's appearance is
`index.rst`, located in a sub-directory of the local `path` named `"docs"` - in
R terms, at `file.path(path, "docs", "index.rst")`.

## Further Information

An RTD site is primarily determined by a few simple files. The
`r2readthedocs()` function initially inserts these files, and then generates
the entire site, creating a `"docs"` sub-directory with a large number of
files. The `rtd_clean()` function can be used to remove all files automatically
generated by RTD itself, reducing the files down to the primary set of files
controlling the site's structure and appearance. These are:

1. `index.rst` responsible for the structure of the main page of the site.
2. `conf.py` A python script for configuring aspects of the site.
3. `requirements.txt` containing a list of required python packages, which may
   be extended as desired.
4. `make.bat` and `Makefile` which are responsible for the main build system,
   and should **not** be modified. Type `make` in the local `"docs"` directory
   to see a full list of `make` options.

The main RTD documentation files can be re-generated at any time with
`rtd_build()`.
